---
title: "Advent of Code Day 02"
author: "Lizzie Pearmain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  github_document:
    toc: true
    toc_depth: 2
editor_options: 
  chunk_output_type: console
---

```{r setup, include = F}
library(dplyr)
library(stringr)
library(knitr)
library(kableExtra)
```

# Part 1

Format: '1-3 a: abcde' means the password must contain between 1 and 3 instances of the letter 'a', and the string after the colon is the password. I have to check whether each password obeys its rule.


```{r read-input}

## mini test set
test.v <- c("1-3 a: abcde", "1-3 b: cdefg", "2-9 c: ccccccccc")
test.df <- data.frame(string = test.v)
glimpse(test.df)

## full set
v <- readLines("input.txt")
df <- data.frame(string = v)
glimpse(df)
```


## make rule

Function: Split string; define the rule; assess whether the password meets the rule.

```{r get-rule}

## define function
assess_rule <- function(string) {
  
  splits <- as.character(unlist(strsplit(string, "-|\\s|:")))
  splits <- splits[splits != ""]  # remove empty strings
  
  r.min <- as.numeric(splits[1])
  r.max <- as.numeric(splits[2])
  r.letter <- as.character(splits[3])
  r.passwd <- as.character(splits[4])
  
  r.n <- str_count(r.passwd, r.letter)
  
  r.pass <- (r.n >= r.min & r.n <= r.max)

  return(r.pass)
  
}

## Vectorize the function
assess_rule_v <- Vectorize(assess_rule, vectorize.args = "string")

## test on mini set
assess_rule_v(test.v)

```


## apply rule

Apply to `string` field in data frame using `dplyr::mutate()`

```{r apply-rule}

df <- df %>%
  mutate(passes = assess_rule_v(string))

glimpse(df)

```


## get answer

How many passwords are valid?

```{r get-ans}

df %>%
  group_by(passes) %>%
  count() %>%
  kable()

```



***



# Part 2 

New rule: The numbers actually refer to positions in the password (indexing from 1), and exactly one of those positions must have this letter.

1-3 a: abcde is valid: position 1 contains a and position 3 does not.

## make rule

```{r make-rule-2}

string <- v[1]

## define function
assess_rule_2 <- function(string) {
  
  splits <- as.character(unlist(strsplit(string, "-|\\s|:")))
  splits <- splits[splits != ""]  # remove empty strings
  
  r.a <- as.numeric(splits[1])
  r.b <- as.numeric(splits[2])
  r.letter <- as.character(splits[3])
  r.passwd <- as.character(splits[4])
  
  ## get indices of letter in password
  r.indices <- as.numeric(unlist(gregexpr(pattern = r.letter, r.passwd)))
  
  ## rule: exclusive or statement
  r.pass <- xor((r.a %in% r.indices), (r.b %in% r.indices))

  return(r.pass)
  
}

## Vectorize the function
assess_rule_2_v <- Vectorize(assess_rule_2, vectorize.args = "string")

## test on mini set
assess_rule_2_v(test.v)

```


## apply rule

Apply to `string` field in data frame using `dplyr::mutate()`

```{r apply-rule-2}

df <- df %>%
  mutate(passes_2 = assess_rule_2_v(string))

glimpse(df)

```


## get answer

How many passwords are valid under the new rule?

```{r get-ans-2}

df %>%
  group_by(passes_2) %>%
  count() %>%
  kable()

```

